Distinguishing NFW and Isothermal Density Profiles with Gravitational Lensing

### 2018-04-17 paper.tex
paper - added things in terms of angles

\begin{abstract}

\end{abstract}

@@

Projected surface density at radius R
\begin{equation}
    \Sigma(R) = \int_{-\infty}^{\infty}{\rho(\sqrt{R^2 + z^2}) dz}
\end{equation}

Average projected surface density within radius R
\begin{equation}
    \overline{\Sigma}(R) = \frac{1}{\pi R^2} \int_0^{2\pi}{d\phi \int_0^{R}{dR'~\Sigma(R')R'}}
\end{equation}

Critical surface density
\begin{equation}
    \Sigma_\mathrm{crit} = \frac{c^2}{4\pi G} \frac{D_S}{D_{SL} D_L}
\end{equation}

Convergence at radius R
\begin{equation}
    \kappa(R) = \frac{\Sigma(R)}{\Sigma_\mathrm{crit}}
\end{equation}

Tangential shear at radius R
\begin{equation}
    \gamma_t(R) = \overline{\kappa}(R) - \kappa(R)
\end{equation}

\[\gamma_1 = -\gamma_t \cos{2\phi}\]
\[\gamma_2 = -\gamma_t \sin{2\phi}\]

Deflection angle
\begin{equation}
    \vec{\alpha}(\vec{\theta}) = \overline{\kappa}(\theta)\vec{\theta}
\end{equation}

Ellipticity
\begin{equation}
    \epsilon(R) = \frac{2 \gamma/(1 - \kappa)}{1 + (\gamma/(1 - \kappa))^2}
\end{equation}

@@

\newpage

@@

\begin{equation}
    \rho_\mathrm{NFW}(r) = \frac{\rho_\mathrm{crit} \delta_c}{(r/r_s)\left(1 + r/r_s\right)^2}
\end{equation}

\begin{equation}
    \Sigma_\mathrm{NFW}(R) = \frac{2 \rho_\mathrm{crit} \delta_c r_s}{(R/r_s)^2 - 1} \left(1 - \frac{2}{\sqrt{(R/r_s)^2 - 1}} \right) \arctan\left(\sqrt{\frac{R/r_s - 1}{R/r_s + 1}} \right)
\end{equation}
[Bartelmann 2001]

\begin{equation}
    \overline{\Sigma}_\mathrm{NFW}(R) = \frac{4 \rho_\mathrm{crit} \delta_c r_s}{(R/r_s)^2} \left(
        \frac{2}{\sqrt{(R/r_s)^2 - 1}} ~\arctan\left(\sqrt{\frac{R/r_s - 1}{R/r_s + 1}} \right) + \ln{\left(\frac{R/r_s}{2}\right)}
    \right)
\end{equation}

\begin{equation}
    \gamma(R) =
\end{equation}

We switch the dependence to $M_{200}$ and $c$ with:

\begin{equation}
    \delta_c = \frac{200}{3} \frac{c^3}{\ln(1 + c) - c/(1 + c)}
\end{equation}

\begin{equation}
    r_s = \frac{1}{c} \left(\frac{3M_{200}}{800\pi \rho_\mathrm{crit}}\right)^\frac{1}{3}
\end{equation}

\begin{equation}
    \rho_\mathrm{iso}(r) = \frac{\sigma^2}{2\pi G (r^2 + {r_c}^2)}
\end{equation}

\begin{equation}
    \Sigma_\mathrm{iso}(R) = \frac{\sigma^2}{2 G \sqrt{R^2 + {r_c}^2}}
\end{equation}

\begin{equation}
    \overline{\Sigma}_\mathrm{iso}(R) = \frac{\sigma^2 \left(\sqrt{R^2 + {r_c}^2} - r_c \right)}{G R^2}
\end{equation}

\begin{equation}
    \gamma(R) = \frac{\sigma^2 \left(\sqrt{R^2 + {r_c}^2} - r_c \right)}{\Sigma_\mathrm{crit} G R^2} - \frac{\sigma^2}{2 \Sigma_\mathrm{crit} G \sqrt{R^2 + {r_c}^2}}
\end{equation}

\begin{equation}
    r_{200} = \left( \frac{3 M_{200}}{800 \pi \rho_\mathrm{crit}} \right)^{1/3}
\end{equation}

\begin{equation}
    \sigma^2 = \frac{G M_{200}}{2 \left(
        \frac{3 M_{200}}{800 \pi \rho_\mathrm{crit}}
        - r_c \arctan \left(
            \frac{3 M_{200}}{800 \pi \rho_\mathrm{crit} {r_c}^3}
        \right)
    \right)}
\end{equation}

$M_{200}$ is defined as:
\begin{equation}
    M_{200} = M_{enc}(r_{200}) = 200 \rho_\mathrm{crit} \frac{4}{3} \pi r_{200}^3
\end{equation}

@@

N sets of e1, e2, theta1, theta2

@@

50 gals/square arcminute
5 arcminutes
$z_L = 0.3$
$z_S = 1$
$M_halo = 10^15$ solar masses
instrinsic ellipticity from gaussian with width 0.2

>>>>

\begin{abstract}
We attempt to distinguish NFW and cored isothermal density profiles using weak gravitational lensing shear.
\end{abstract}

@@

Projected surface density at radius R
\[\Sigma(R) = \int_{-\infty}^{\infty}{\rho(\sqrt{R^2 + z^2}) dz}\]

Average projected surface density within radius R
\[\overline{\Sigma}(R) = \frac{1}{\pi R^2} \int_0^{2\pi}{d\phi \int_0^{R}{dR'~\Sigma(R')R'}}\]

Critical surface density
\[\Sigma_\mathrm{crit} = \frac{c^2}{4\pi G} \frac{D_S}{D_{SL} D_L}\]

Convergence
\[\kappa(\theta) = \frac{\Sigma(\theta)}{\Sigma_\mathrm{crit}}\]

Tangential shear
\[\gamma_t(\theta) = \overline{\kappa}(\theta) - \kappa(\theta)\]
\[\gamma_1 = -\gamma_t \cos{2\phi}\]
\[\gamma_2 = -\gamma_t \sin{2\phi}\]
\[\gamma = \gamma_t = \sqrt{{\gamma_1}^2 + {\gamma_2}^2} = -\gamma_1 \cos{2\phi} -\gamma_2 \sin{2\phi}\]

Deflection angle
\[\vec{\alpha}(\vec{\theta}) = \overline{\kappa}(\theta)\vec{\theta}\]
\[\vec{\beta} = \vec{\theta} - \vec{\alpha} = (1 - \overline{\kappa}(\theta))\vec{\theta} \]

Ellipticity
\[\epsilon_i = \frac{2 \gamma_i/(1 - \kappa)}{1 + \gamma^2/(1 - \kappa)^2}\]

In small angle approximation, any length $R = D_L \theta$.

@@

\[\rho_\mathrm{iso}(r) = \frac{\sigma^2}{2\pi G (r^2 + {r_c}^2)}\]

\[\Sigma_\mathrm{iso}(R) = \frac{\sigma^2}{2 G \sqrt{R^2 + {r_c}^2}}\]

\[\overline{\Sigma}_\mathrm{iso}(R) = \frac{\sigma^2 \left(\sqrt{R^2 + {r_c}^2} - r_c \right)}{G R^2}\]

In terms of angles:

\[\Sigma_\mathrm{iso}(\theta) = \frac{\sigma^2}{2 G D_L \sqrt{\theta^2 + {\theta_c}^2}}\]

\[\overline{\Sigma}_\mathrm{iso}(\theta) = \frac{\sigma^2 \left(\sqrt{\theta^2 + {\theta_c}^2} - \theta_c \right)}{G D_L \theta^2}\]

\[\gamma_\mathrm{iso}(\theta) = \frac{\sigma^2 \left(\sqrt{\theta^2 + {\theta_c}^2} - \theta_c \right)}{\Sigma_\mathrm{crit} G D_L \theta^2} - \frac{\sigma^2}{2 \Sigma_\mathrm{crit} G D_L \sqrt{\theta^2 + {\theta_c}^2}}\]

\[\sigma^2 = \frac{G M_{200}}{2 \left(
    \frac{3 M_{200}}{800 \pi \rho_\mathrm{crit}}
    - r_c \arctan \left(
        \frac{3 M_{200}}{800 \pi \rho_\mathrm{crit} {r_c}^3}
    \right)
\right)}\]
\[\sigma^2 = \frac{G M_{200}}{2 \left(
    \frac{3 M_{200}}{800 \pi \rho_\mathrm{crit}}
    - D_L \theta_c \arctan \left(
        \frac{3 M_{200}}{800 \pi \rho_\mathrm{crit} {D_L}^3 {\theta_c}^3}
    \right)
\right)}\]

This is derived from the definition of $M_{200}$:

\[M_{200} = M_{enc}(r_{200}) = 200 \rho_\mathrm{crit} \frac{4}{3} \pi {r_{200}}^3\]
\[r_{200} = \left( \frac{3 M_{200}}{800 \pi \rho_\mathrm{crit}} \right)^{1/3}\]

Ellipticity equations are very ugly but trivial to calculate from the shear.

@@

\[\rho_\mathrm{NFW}(r) = \frac{\rho_\mathrm{crit} \delta_C}{(r/r_s)\left(1 + r/r_s\right)^2}\]

\[\Sigma_\mathrm{NFW}(R) = \frac{2 \rho_\mathrm{crit} \delta_C r_s}{(R/r_s)^2 - 1} \left(1 - \frac{2}{\sqrt{(R/r_s)^2 - 1}} \right) \arctan\left(\sqrt{\frac{R/r_s - 1}{R/r_s + 1}} \right)\]
[Bartelmann 2001]

\[\overline{\Sigma}_\mathrm{NFW}(R) = \frac{4 \rho_\mathrm{crit} \delta_C r_s}{(R/r_s)^2} \left(
    \frac{2}{\sqrt{(R/r_s)^2 - 1}} ~\arctan\left(\sqrt{\frac{R/r_s - 1}{R/r_s + 1}} \right) + \ln{\left(\frac{R/r_s}{2}\right)}
\right)\]

In terms of angles:

\[\Sigma_\mathrm{NFW}(\theta) = \frac{2 \rho_\mathrm{crit} \delta_C D_L \theta_s}{(\theta/\theta_s)^2 - 1} \left(1 - \frac{2}{\sqrt{(\theta/\theta_s)^2 - 1}} \right) \arctan\left(\sqrt{\frac{\theta/\theta_s - 1}{\theta/\theta_s + 1}} \right)\]

\[\overline{\Sigma}_\mathrm{NFW}(\theta) = \frac{4 \rho_\mathrm{crit} \delta_C D_L \theta_s}{(\theta/\theta_s)^2} \left(
    \frac{2}{\sqrt{(\theta/\theta_s)^2 - 1}} ~\arctan\left(\sqrt{\frac{\theta/\theta_s - 1}{\theta/\theta_s + 1}} \right) + \ln{\left(\frac{\theta/\theta_s}{2}\right)}
\right)\]

\[\gamma_\mathrm{NFW}(\theta) = \frac{\overline{\Sigma}_\mathrm{NFW}(\theta) - \Sigma_\mathrm{NFW}(\theta)}{\Sigma_\mathrm{crit}} \]

Can calculate ellipticities from tangential shear.

We switch the dependence to $M_{200}$ and $C$ with:

\[\delta_C = \frac{200}{3} \frac{C^3}{\ln(1 + C) - C/(1 + C)}\]

\[r_s = \frac{1}{C} \left(\frac{3M_{200}}{800\pi \rho_\mathrm{crit}}\right)^\frac{1}{3}\]

\[C = \frac{r_{200}}{r_s}\]

@@

Fake data:
N sets of e1, e2, theta1, theta2

@@

50 gals/square arcminute
5 arcminutes
$z_L = 0.3$
$z_S = 1$
$M_{halo} = 10^15$ solar masses
instrinsic ellipticity from gaussian with width 0.2
